---
title: "FA_AutomobileCompanies"
author: "Kajori De"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  markdown:
    wrap: 75
  html_document:
    df_print: paged
---

```{r echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, warning = FALSE)
```

## Introduction

This report analyzes the financial performance of four publicly traded
companies: Ford (F), General Motors (GM), Tesla (TSLA), and Toyota Motors
(TM), over the past year. The analysis focuses on daily stock returns, risk
measures, and comparisons to the S&P 500 Index (\^GSPC).

## Load packages

Packages - The necessary libraries are loaded, including quantmod for
retrieving financial data, ggplot2 for data visualization, reshape2 and
dplyr for data manipulation, zoo for time series data, plotly for
interactive plots, knitr and rmarkdown for report generation.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(quantmod)
library(ggplot2)
library(reshape2)
library(dplyr)
library(zoo)
library(plotly)
library(hrbrthemes)
```

## Data Retrieval

Financial data for the specified companies and the S&P 500 index is
retrieved from Yahoo Finance using the `quantmod` package. The data covers
a one-year period ending on the current date.

```{r}
# Define companies to analyze
companies <- c("F", "GM", "TSLA", "TM")
start_date <- Sys.Date() - 365
end_date <- Sys.Date()

# Retrieve data for companies and market index
getSymbols(companies, src = "yahoo", from = start_date, to = end_date)
getSymbols("^GSPC", src = "yahoo", from = start_date, to = end_date)
```

## Financial Analysis

### Daily Returns

-   Daily returns represent the percentage change in a stock's price from
    one day to the next. Positive values indicate a price increase, while
    negative values indicate a decrease.

-   They are calculated using the logarithmic difference of consecutive
    adjusted closing prices (`diff(log(stock_data))`). This approach
    provides a percentage change rather than the absolute difference. Each
    company's daily return is stored in a separate data frame named after
    the company symbol with the suffix `.Ret`.

-   A 7-day rolling mean is applied to each company's return series using
    `rollapply` to smooth out short-term fluctuations and highlight
    underlying trends. Smoothing the data with a rolling mean helps
    visualize the general trend in returns and reduces the impact of daily
    market noise.

```{r}
# Calculate daily returns and handle missing data
for (company in companies) {
  stock_data <- get(company)
  stock_data <- na.omit(stock_data)
  stock_data <- stock_data[, paste0(company, ".Adjusted")]
  stock_data <- diff(log(stock_data))
  colnames(stock_data) <- paste0(company, ".Ret")
  assign(company, stock_data)
}

# Calculate S&P 500 returns
GSPC <- na.omit(GSPC)
GSPC <- GSPC[, "GSPC.Adjusted"]
GSPC <- diff(log(GSPC))
colnames(GSPC) <- "GSPC.Ret"

# Combine returns into a single data frame
returns_data <- na.omit(merge(F, GM, TSLA, TM, GSPC))

# Convert to data.frame for manipulation
returns_data_df <- data.frame(Date = index(returns_data), coredata(returns_data))

# Apply smoothing (7-day rolling mean)
returns_data_smooth <- returns_data_df %>%
  mutate(across(-Date, ~ rollapply(.x, width = 7, FUN = mean, fill = NA, align = "right")))

# Convert back to xts
returns_data_smooth_xts <- xts(returns_data_smooth[, -1], order.by = returns_data_smooth$Date)

# Interactive Smoothed Daily Returns

returns_long_smooth <- as.data.frame(returns_data_smooth_xts)
returns_long_smooth$Date <- index(returns_data_smooth_xts)
returns_long_smooth <- melt(returns_long_smooth, id.vars = "Date")

dailyreturns_graph <- ggplot(returns_long_smooth, aes(x = Date, y = value, color = variable)) +
  geom_line() +
  labs(
    title = "Smoothed Daily Returns (7-day Rolling Mean)",
    x = "Date", y = "Return", color = "Company"
  ) +
  theme_ipsum()

ggplotly(dailyreturns_graph)
```

### Capital Asset Pricing Model (CAPM)

The CAPM estimates the expected return of an asset based on its systematic
risk (beta) relative to the market. A CAPM regression is performed for each
company using the `lm` function. The regression model is: Company Return \~
Market Return, where the market return is represented by the S&P 500 daily
return (`GSPC.Ret`).

```{r}
capm_results <- list()
for (company in companies) {
  formula <- as.formula(paste0(company, ".Ret ~ GSPC.Ret"))
  capm_model <- lm(formula, data = returns_data)
  capm_results[[company]] <- summary(capm_model)
}
```

### Beta Coefficient

Beta indicates how much a company's returns move relative to the market
(S&P 500). A beta of 1 suggests the company's returns mirror the market,
while a beta greater than 1 indicates higher volatility compared to the
market.

```{r}
beta_coefficient <- sapply(companies, function(company) {
  capm_results[[company]]$coefficients["GSPC.Ret", "Estimate"]
})
```

### Sharpe Ratio

The Sharpe Ratio is a risk-adjusted performance measure. It calculates the
excess return (return above the risk-free rate) per unit of risk (standard
deviation of return). A higher Sharpe Ratio indicates better risk-adjusted
performance. I assume a risk-free rate of 1% per year, converted to daily
basis (0.01 / 252).

```{r}
# Sharpe Ratios
risk_free_rate <- 0.01 / 252 # Assume 1% annual risk-free rate, converted to daily
sharpe_ratios <- sapply(companies, function(company) {
  column_name <- paste0(company, ".Ret")
  mean_return <- mean(returns_data[, column_name], na.rm = TRUE)
  sd_return <- sd(returns_data[, column_name], na.rm = TRUE)
  (mean_return - risk_free_rate) / sd_return
})
```

### Interpreting Beta Coefficient and Sharpe Ratio

The beta_sharpe_table summarizes the beta coefficient and sharpe ratio for
each company.

Tesla offers the potential for higher returns but comes with significantly
more risk. Toyota Motors appears to be the most conservative option with
lower volatility but also potentially lower returns. General Motors and
Ford seem to fall somewhere in between, with General Motors being closer to
the market's volatility and F being a bit more volatile.

[Caveat]{.underline} - This analysis is based on a limited timeframe (one
year) and specific assumptions (risk-free rate). Other factors beyond beta
and Sharpe Ratio can influence investment decisions.

```{r}
beta_sharpe_table <- data.frame(
  Company = companies,
  Beta_Coefficient = round(beta_coefficient, 2),
  Sharpe_Ratio = round(sharpe_ratios, 2)
)

beta_sharpe_table


beta_sharpe_graph <- ggplot(beta_sharpe_table, aes(x = Company)) +
  geom_bar(aes(y = Beta_Coefficient),
    stat = "identity", fill = "cornsilk3", alpha = 0.7
  ) +
  geom_point(aes(y = Sharpe_Ratio * 10),
    color = "darkcyan", size = 4
  ) +
  labs(
    title = "Beta Coefficients and Sharpe Ratios",
    x = "Company", y = "Value"
  ) +
  scale_y_continuous(sec.axis = sec_axis(~ . / 10, name = "Sharpe Ratio")) +
  theme_ipsum()

beta_sharpe_graph
```

### Security Market Line (SML)

The SML depicts the expected relationship between a stock's beta
coefficient and its Sharpe Ratio. Its slope represents the market risk
premium, which is the additional return investors expect for taking on more
risk (higher beta) compared to the risk-free rate. Stocks plotting above
the SML are considered to have potentially better risk-adjusted returns,
while those below may be less attractive.

As can be seen in the graph, Toyota and General Motors better risk-adjusted
returns while Tesla and Ford do not.

```{r}
# Security Market Line (SML)
sml_plot <- ggplot(data = beta_sharpe_table, aes(x = Beta_Coefficient, y = Sharpe_Ratio, color = Company)) +
  geom_point(size = 4) +
  geom_abline(
    intercept = 0, slope = max(beta_sharpe_table$Sharpe_Ratio) / max(beta_sharpe_table$Beta_Coefficient),
    color = "navyblue"
  ) +
  labs(
    title = "Security Market Line (SML)",
    x = "Beta Coefficient",
    y = "Sharpe Ratio",
    color = "Company"
  ) +
  theme_ipsum()

sml_plot
```

### Hypothesis Testing

The t-test is used to compare the returns of each stock with the S&P 500
index to determine if there is a statistically significant difference.

Companies with p-values below a certain threshold (e.g., 0.05) are
considered to have significantly different returns compared to the market.

**Ford, General Motors, and Tesla** have high p-values (greater than 0.05).
This suggests we **fail to reject the null hypothesis**. In other words,
there is **not enough evidence** to conclude that the average daily returns
of Ford, General Motors, and Tesla are statistically different from the S&P
500's return at the 5% significance level.

Whereas **Toyota** has a higher t-statistic (positive) and a p-value
approaching 0.5. While not statistically significant at the 5% level, the
positive t-statistic might indicate a trend for Toyota's average daily
return to be slightly higher than the S&P 500's return.

[Caveat]{.underline} - While the t-tests don't reveal statistically
significant differences, this doesn't necessarily imply the returns are
identical. There might be smaller differences that aren't statistically
significant with this sample size. These results are based on daily returns
over a one-year period. Looking at longer timeframes or different time
periods could yield different conclusions.

```{r}
# t-tests comparing returns
t_tests <- lapply(companies, function(company) {
  column_name <- paste0(company, ".Ret")
  t_result <- t.test(returns_data[, column_name], returns_data$GSPC.Ret)
  list(
    t_statistic = t_result$statistic,
    p_value = t_result$p.value,
    CI_lower = t_result$conf.int[1],
    CI_upper = t_result$conf.int[2]
  )
})

# data frame from the results
t_tests_df <- do.call(rbind, t_tests)
rownames(t_tests_df) <- companies
colnames(t_tests_df) <- c("t_statistic", "p_value", "CI_lower", "CI_upper")

t_tests_df
```

The box plots visually represent the distribution of daily returns for each
company. The center line within the box represents the median return, while
the box encompasses the middle 50% of the data. The whiskers extend to the
most extreme data points within 1.5 times the interquartile range from the
box. Outliers are shown as individual points beyond the whiskers.

As can be seen in the graph, Tesla shows the widest range of returns,
indicating high volatility. Toyota Motors exhibits the most stable returns,
with the smallest interquartile range. Ford and General Motors have similar
distributions, with some outliers, suggesting moderate volatility compared
to Tesla and Toyota.

```{r}
returns_long <- returns_data_df %>%
  select(-Date) %>%
  pivot_longer(cols = everything(), names_to = "Company", values_to = "Return")

t_tests_plots <- ggplot(
  returns_long,
  aes(x = Company, y = Return, fill = Company)
) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title = "Box Plot of Daily Returns for Each Company",
    x = "Company", y = "Daily Return"
  ) +
  theme_ipsum() +
  theme(legend.position = "none")

t_tests_plots
```

### Value at Risk (VaR)

VaR estimates the potential loss in value of an asset or portfolio over a
defined period for a given confidence interval. In this code, the 1-day VaR
at the 95% confidence level is calculated by determining the quantile of
the daily returns distribution that corresponds to the worst 5% of
outcomes.

The 1-day 95% VaR for Ford is -3.32%, meaning there is a 5% chance that
Ford's stock could lose more than 3.32% of its value in one day. Tesla has
the highest VaR at -4.78%, indicating the greatest potential for daily loss
among the companies analyzed, while Toyota has the lowest VaR at -2.41%,
suggesting the least risk.

```{r}
# VaR calculation function
calculate_var <- function(returns, confidence_level = 0.95) {
  quantile(returns, probs = 1 - confidence_level, na.rm = TRUE)
}

# Calculate VaR for each company
var_95 <- sapply(companies, function(company) {
  column_name <- paste0(company, ".Ret")
  calculate_var(returns_data[, column_name], confidence_level = 0.95)
})

# Remove names from var_95
var_95 <- unname(var_95)

# Create a data frame for VaR results
var_results <- data.frame(
  Company = companies,
  VaR_95 = round(var_95, 4)
)

# Print VaR results
print(var_results)
```

The bar chart below shows the 1-day 95% VaR for each company. Here, a
higher bar signifies greater potential daily downside risk. Tesla has the
highest VaR, suggesting the most significant potential loss, while Toyota
exhibits the lowest VaR, implying the least risk.

```{r}
var_plot <- ggplot(var_results, aes(x = Company, y = VaR_95, fill = Company)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Value at Risk (VaR) at 95% Confidence Level",
    x = "Company", y = "VaR"
  ) +
  theme_ipsum()

var_plot
```

## Assumptions and Limitations

-   This analysis is based on historical data over a one-year period and
    assumes the accuracy of the data retrieved from Yahoo Finance.

-   The risk-free rate is assumed to be 1% per year for illustrative
    purposes.

-   The chosen financial metrics and tests provide insights but may not
    capture all aspects of investment risk.

-   The results may not be generalizable to different timeframes or
    economic conditions.
